---
title: "NIFA clustering no incorporation"
author: "Cassandra Wattenburger"
date: "5/8/2023"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
```

```{r}
# Clear working directory, load in packages, generate package info
rm(list=ls())

library("tidyverse")

sessionInfo()
```

# Import data

```{r}
# Import growth data
growth <- readRDS("../data_intermediate/NIFA_grests_final_ind.rds")

# Paprica estimates
pap <- readRDS("../data_intermediate/NIFA_paprica_ests.rds")

# Incorporation
subnum <- readRDS("../data_intermediate/NIFA_incorp_profiles.rds") %>% 
  select(Soil, ASV, substrate_num)

# Combine
growth_pap_subnum <- inner_join(growth, pap) %>% 
  mutate(length = end_day - start_day, # add length of growth feature
         start_abund_corr = start_abund/n16S, # correct for n16S
         end_abund_corr = end_abund/n16S,
         change_abund_corr = change_abund/n16S) %>% 
  inner_join(subnum) %>% 
  select(label, Soil, Replicate, ASV, g, start_day, end_day, length, start_abund_corr, end_abund_corr, change_abund_corr, n16S, genome_size, npaths_actual, nec_actual, substrate_num)
growth_pap_subnum
```

Average for each ASV:

```{r}
# Growth metrics ASV averages, across soils
growth_asv <- growth_pap_subnum %>% 
  group_by(Soil, ASV) %>%
  summarize(g = mean(g),
            start_day = mean(start_day),
            end_day = mean(end_day),
            length = mean(length),
            start_abund_corr = mean(start_abund_corr),
            end_abund_corr = mean(end_abund_corr),
            change_abund_corr = mean(change_abund_corr),
            substrate_num = mean(substrate_num)) %>% 
  ungroup()
growth_asv

features_ag <- growth_asv %>% 
  filter(Soil=="Ag")
features_ag

features_meadow <- growth_asv %>% 
  filter(Soil=="Meadow")
features_meadow
```

# Preparation

## Check for co-correlations

Check for co-correlations between variables:

```{r}
cor(features_ag[, c('g', 'start_day', "end_day", "length", 'start_abund_corr', 'end_abund_corr', 'change_abund_corr', 'substrate_num')])

cor(features_meadow[, c('g', 'start_day', "end_day", "length", 'start_abund_corr', 'end_abund_corr', 'change_abund_corr', 'substrate_num')])
```

Remove highly co-correlated features and paprica estimates:

```{r}
features_ag2 <- select(features_ag, ASV, g:length, start_abund_corr, change_abund_corr,
                       substrate_num)
features_meadow2 <- select(features_meadow, ASV, g:length, start_abund_corr, change_abund_corr,
                           substrate_num)
```

Visualize feature distrubutions:

```{r}
features_ag2_long <- features_ag2 %>%  
  pivot_longer(cols=c(g:substrate_num), names_to="feature", values_to="value")
features_ag2_long

features_meadow2_long <- features_meadow2 %>%  
  pivot_longer(cols=c(g:substrate_num), names_to="feature", values_to="value")
features_meadow2_long

# Raw
features_ag2_long %>% 
  ggplot(aes(x=value)) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()

features_meadow2_long %>% 
  ggplot(aes(x=value)) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()

# Log
features_ag2_long %>% 
  ggplot(aes(x=log(value))) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()

features_meadow2_long %>% 
  ggplot(aes(x=log(value))) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()
```

## Transformations and scaling

Log transform:

```{r}
features_ag3 <- features_ag2 %>% 
  mutate(start_day = start_day+1) %>% # can't log transform 0, add 1 to all values of start_day
  mutate_if(is.numeric, log)
features_ag3

features_meadow3 <- features_meadow2 %>% 
  mutate(start_day = start_day+1) %>% # can't log transform 0, add 1 to all values of start_day
  mutate_if(is.numeric, log)
features_meadow3
```

Scale:

```{r}
features_ag4 <- features_ag3 %>% 
  mutate_if(is.numeric, scale)
features_ag4

features_meadow4 <- features_meadow3 %>% 
  mutate_if(is.numeric, scale)
features_meadow4
```

## Add labels

```{r}
features_ag5 <- features_ag4 %>%
  column_to_rownames(var="ASV") 
features_ag5

features_meadow5 <- features_meadow4 %>%
  column_to_rownames(var="ASV") 
features_meadow5
```

Re-visualize feature distributions:

```{r}
features_ag5_long <- features_ag5 %>%  
  pivot_longer(cols=c(g:substrate_num), names_to="feature", values_to="value")
features_ag5_long

features_meadow5_long <- features_meadow5 %>%  
  pivot_longer(cols=c(g:substrate_num), names_to="feature", values_to="value")
features_meadow5_long

features_ag5_long %>% 
  ggplot(aes(x=value)) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()

features_meadow5_long %>% 
  ggplot(aes(x=value)) +
  geom_density() +
  facet_wrap(~feature, scales="free") +
  theme_test()
```

# Principle components

```{r}
# Remove NAs if any
features_ag6 <- features_ag5[complete.cases(features_ag5),]
features_meadow6 <- features_meadow5[complete.cases(features_meadow5),]

features_ag6_pcs <- prcomp(features_ag6)
features_meadow6_pcs <- prcomp(features_meadow6)

summary(features_ag6_pcs)
summary(features_meadow6_pcs)

# Create dataframe
features_ag6_pcs_df <- features_ag6_pcs$x %>% 
  data.frame()
features_ag6_pcs_df

features_meadow6_pcs_df <- features_meadow6_pcs$x %>% 
  data.frame()
features_meadow6_pcs_df
```

Visualize:

```{r}
# PC 1 & 2
features_ag6_pcs_df %>% 
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point() +
  theme_test()

features_meadow6_pcs_df %>% 
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point() +
  theme_test()
```

```{r}
# Cumulative proportion of variance explained
summary(features_ag6_pcs)$importance[3,] %>%
  as.data.frame() %>% 
  rownames_to_column(var="PC") %>% 
  rename(cum_prop='.') %>% 
  ggplot(aes(x=PC, y=cum_prop)) +
  geom_point() +
  geom_line() +
  theme_test()

summary(features_meadow6_pcs)$importance[3,] %>%
  as.data.frame() %>% 
  rownames_to_column(var="PC") %>% 
  rename(cum_prop='.') %>% 
  ggplot(aes(x=PC, y=cum_prop)) +
  geom_point() +
  geom_line() +
  theme_test()

# Supp figure
ag6_cumvar <- summary(features_ag6_pcs)$importance[3,] %>%
  as.data.frame() %>% 
  rownames_to_column(var="PC") %>% 
  rename(cum_prop='.') %>% 
  add_column(soil = "agriculture")

meadow6_cumvar <- summary(features_meadow6_pcs)$importance[3,] %>%
  as.data.frame() %>% 
  rownames_to_column(var="PC") %>% 
  rename(cum_prop='.') %>% 
  add_column(soil = "meadow")

suppfig_pca <- ag6_cumvar %>% 
  bind_rows(meadow6_cumvar) %>% 
  ggplot(aes(x=PC, y=cum_prop, shape=soil)) +
  geom_point(size=2) +
  labs(x="principle components", y="cumulative variance explained") +
  theme_test() +
  theme(axis.title = element_text(size=14),
      axis.text = element_text(size=12),
      legend.title = element_text(size=14),
      axis.title.x = element_blank(),
      legend.text = element_text(size=12))
suppfig_pca

# Scree plot
eigenvals_ag <- features_ag6_pcs$sdev^2 %>% 
  data.frame() %>% 
  add_column(PC=1:nrow(.)) %>% 
  rename(eigenvalue=".") %>% 
  mutate(PC = as.factor(PC))
eigenvals_ag

eigenvals_ag %>% 
  ggplot(aes(x=PC, y=eigenvalue)) +
  geom_point() +
  geom_line() +
  theme_test()

eigenvals_meadow <- features_meadow6_pcs$sdev^2 %>% 
  data.frame() %>% 
  add_column(PC=1:nrow(.)) %>% 
  rename(eigenvalue=".") %>% 
  mutate(PC = as.factor(PC))
eigenvals_meadow

eigenvals_meadow %>% 
  ggplot(aes(x=PC, y=eigenvalue)) +
  geom_point() +
  geom_line() +
  theme_test()
```

Shoulder is at 3-4 PCs.

```{r, eval=FALSE}
ggsave(suppfig_pca, file="../figures/suppfig_pca.svg", units="mm", height=90, width=150, device="svg")
```


# Within-sum squared distances

```{r}
# testing diff nums clusters 1-6
library(cluster)
set.seed(123)

# Cluster and extract total within-cluster summed squared distance
km1_ag <- kmeans(features_ag6, centers = 1, nstart = 25)$tot.withinss
km2_ag <- kmeans(features_ag6, centers = 2, nstart = 25)$tot.withinss
km3_ag <- kmeans(features_ag6, centers = 3, nstart = 25)$tot.withinss
km4_ag <- kmeans(features_ag6, centers = 4, nstart = 25)$tot.withinss
km5_ag <- kmeans(features_ag6, centers = 5, nstart = 25)$tot.withinss
km6_ag <- kmeans(features_ag6, centers = 6, nstart = 25)$tot.withinss

km1_meadow <- kmeans(features_meadow6, centers = 1, nstart = 25)$tot.withinss
km2_meadow <- kmeans(features_meadow6, centers = 2, nstart = 25)$tot.withinss
km3_meadow <- kmeans(features_meadow6, centers = 3, nstart = 25)$tot.withinss
km4_meadow <- kmeans(features_meadow6, centers = 4, nstart = 25)$tot.withinss
km5_meadow <- kmeans(features_meadow6, centers = 5, nstart = 25)$tot.withinss
km6_meadow <- kmeans(features_meadow6, centers = 6, nstart = 25)$tot.withinss

# Create df
ag <- c(km1_ag, km2_ag, km3_ag, km4_ag, km5_ag, km6_ag)
meadow <- c(km1_meadow, km2_meadow, km3_meadow, km4_meadow, km5_meadow, km6_meadow)
totwss_df <- tibble(ag, meadow) %>% 
  add_column(cluster_num = c(1:6)) %>% 
  pivot_longer(c("ag", "meadow"), names_to = "soil")

# Plot
totwss_suppfig <- totwss_df %>% 
  ggplot(aes(x=cluster_num, y=value, shape=soil)) +
  geom_point(size=2) +
  theme_test() +
  labs(y="Total within-cluster squared distance", x="Number of clusters") +
  theme_test() +
  theme(axis.title = element_text(size=14),
      axis.text = element_text(size=12),
      legend.title = element_text(size=14),
      legend.text = element_text(size=12))
totwss_suppfig
```


```{r, eval=FALSE}
ggsave(totwss_suppfig, file="../figures/suppfig_totwss.svg", units="mm", height=90, width=150, device="svg")
```

Elbow clearly at 3 clusters.

# Cluster

K-means:

Choosing n=3 because skree plot above, want to do comparison to CSR triangle theory and previous work in lab with data from same soils, data is a single cloud anyway to cluster choice is somewhat arbritray, using as a basis for looking at how growth metrics tend to self-organize within community.

```{r}
# Import library
library(cluster)

# Cluster
set.seed(123)
km3_ag <- kmeans(features_ag6, centers = 3, nstart = 25)
km3_ag

km3_meadow <- kmeans(features_meadow6, centers = 3, nstart = 25)
km3_meadow

# Cluster key
clusterkey_ag <- km3_ag$cluster %>% 
  data.frame() %>% 
  rename(cluster='.') %>% 
  rownames_to_column(var="ASV") %>% 
  select(ASV, cluster) %>% 
  mutate(cluster = as.character(cluster))
clusterkey_ag

clusterkey_meadow <- km3_meadow$cluster %>% 
  data.frame() %>% 
  rename(cluster='.') %>% 
  rownames_to_column(var="ASV") %>% 
  select(ASV, cluster) %>% 
  mutate(cluster = as.character(cluster)) %>% 
  mutate(cluster = case_when(cluster=="1" ~ "3", # rename clusters to match ag, based on growth metric profiles
                             cluster=="2" ~ "2",
                             cluster=="3" ~ "1"))
clusterkey_meadow
```

Visualize:

On top of first 3 principle components

```{r}
# Add cluster designations
features_clust_ag <- features_ag6_pcs_df %>% 
  rownames_to_column(var="ASV") %>% 
  inner_join(clusterkey_ag)

features_clust_meadow <- features_meadow6_pcs_df %>% 
  rownames_to_column(var="ASV") %>% 
  inner_join(clusterkey_meadow)

features_clust_ag %>% 
  ggplot(aes(x=PC1, y=PC2, color=cluster)) +
  scale_color_manual(values=c("#c75d65", "#929644", "#9473c4")) +
  geom_point() +
  theme_test()

features_clust_meadow %>% 
  ggplot(aes(x=PC1, y=PC2, color=cluster)) +
  scale_color_manual(values=c("#c75d65", "#929644", "#9473c4")) +
  geom_point() +
  theme_test()

# 3D
library(plotly)
pcs_clust_ag <- plot_ly(features_clust_ag, x=~PC1, y=~PC2, z=~PC3, 
                     color=~cluster, marker=list(size=10, line=list(color = 'rgba(0, 0, 0, .8)', width = 2)))
pcs_clust_ag

pcs_clust_meadow <- plot_ly(features_clust_meadow, x=~PC1, y=~PC2, z=~PC3, 
                     color=~cluster, marker=list(size=10, line=list(color = 'rgba(0, 0, 0, .8)', width = 2)))
pcs_clust_meadow
```

Combined:

```{r}
# Combine
features_clust_meadow2 <- features_clust_meadow %>% 
  add_column(Soil="Meadow")

features_clust_ag2 <- features_clust_ag %>% 
  add_column(Soil="Ag")

features_clust <- bind_rows(features_clust_ag2, features_clust_meadow2)

# Visualize
pca_plot <- features_clust %>% 
  mutate(Cluster = case_when(cluster=="1" ~ "Competitor",
                             cluster=="2" ~ "Scarcity",
                             cluster=="3" ~ "Ruderal"),
         Cluster = as_factor(Cluster),
         Cluster = fct_relevel(Cluster, "Competitor", "Scarcity", "Ruderal"),
         Soil = if_else(Soil=="Ag", "Agriculture", "Meadow")) %>% 
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(color=Cluster, shape=Soil), size=2) +
  scale_color_manual(values=c("#9473c4", "#929644", "#c75d65")) +
  scale_shape_manual(values=c(1,16)) +
  theme_test() +
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.text = element_text(size=12))
pca_plot

# without legend
pca_plot2 <- features_clust %>% 
  mutate(Cluster = case_when(cluster=="1" ~ "Competitor",
                             cluster=="2" ~ "Scarcity",
                             cluster=="3" ~ "Ruderal"),
         Cluster = as_factor(Cluster),
         Cluster = fct_relevel(Cluster, "Competitor", "Scarcity", "Ruderal"),
         Soil = if_else(Soil=="Ag", "Agriculture", "Meadow")) %>% 
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(color=Cluster, shape=Soil), size=2) +
  scale_color_manual(values=c("#9473c4", "#929644", "#c75d65")) +
  scale_shape_manual(values=c(1,16)) +
  theme_test() +
  theme(axis.title = element_text(size=14),
        axis.text = element_text(size=12),
        legend.position = "none",
        legend.title =element_blank(),
        legend.text = element_blank())
pca_plot2

# ISME specifications
pca_plot_isme_legend <- features_clust %>% 
  mutate(Cluster = case_when(cluster=="1" ~ "Competitor",
                             cluster=="2" ~ "Scarcity",
                             cluster=="3" ~ "Ruderal"),
         Cluster = as_factor(Cluster),
         Cluster = fct_relevel(Cluster, "Competitor", "Scarcity", "Ruderal"),
         Soil = if_else(Soil=="Ag", "Agriculture", "Meadow")) %>% 
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(color=Cluster, shape=Soil), size=2) +
  scale_color_manual(values=c("#9473c4", "#929644", "#c75d65")) +
  scale_shape_manual(values=c(1,16)) +
  theme_test() +
  theme(axis.title = element_text(size=7),
        axis.text = element_text(size=7),
        legend.title = element_text(size=7),
        legend.text = element_text(size=7))
pca_plot_isme_legend

pca_plot_isme_nolegend <- features_clust %>% 
  mutate(Cluster = case_when(cluster=="1" ~ "Competitor",
                             cluster=="2" ~ "Scarcity",
                             cluster=="3" ~ "Ruderal"),
         Cluster = as_factor(Cluster),
         Cluster = fct_relevel(Cluster, "Competitor", "Scarcity", "Ruderal"),
         Soil = if_else(Soil=="Ag", "Agriculture", "Meadow")) %>% 
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(color=Cluster, shape=Soil), size=2) +
  scale_color_manual(values=c("#9473c4", "#929644", "#c75d65")) +
  scale_shape_manual(values=c(1,16)) +
  theme_test() +
  theme(axis.title = element_text(size=7),
        axis.text = element_text(size=7),
        legend.position = "none")
pca_plot_isme_nolegend
```

Export figure:

```{r,  eval=FALSE}
ggsave(pca_plot_isme_legend, filename="../figures/fig_pcaclusters_isme_legend.svg", units="mm", width=85, height=85, device="svg")
ggsave(pca_plot_isme_nolegend, filename="../figures/fig_pcaclusters_isme_nolegend.svg", units="mm", width=85, height=85, device="svg")
```

Save combined:

```{r,e val=FALSE}
cluster_key_all <- select(features_clust, Soil, ASV, cluster)
saveRDS(cluster_key_all, file="../data_intermediate/NIFA_cluster_sepsoil_subnum.rds")
```

