---
title: "SFA2 - prerocessing"
author: "Cassandra Wattenburger"
date: "9/15/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
```

```{r}
# Clear working directory, load in packages, generate package info
rm(list=ls())

library("tidyverse")
library("ape")
library("phyloseq")

sessionInfo()
```

# Import data

* counts
* taxonomy
* tree
* meta

```{r}
count <- read_tsv(file="../data_amplicon/NIFA2/final/NIFA2.counts-final.tsv")
tax <- read_tsv(file="../data_amplicon/NIFA2/final/NIFA2.taxonomy-final.tsv")
tree <- read.tree(file="~/SFA/data_amplicon/SFA2_full/tree/SFA2_full.tree-final.nwk")
meta <- read_tsv("../NIFA_metadata.tsv") %>% 
  mutate(SampleID = as.character(SampleID))
```

# Import into phyloseq

```{r}
# Reformat and import count table
count_matrix <- count %>%
  column_to_rownames(var="ASV") %>%
  rename_with(~ gsub("(.+)", "sa\\1", .x)) %>%
  as.matrix()

phy_count <- otu_table(count_matrix, taxa_are_rows=TRUE)

# Reformat and import taxonomy table
tax_matrix <- tax %>%
  mutate(dummy=ASV) %>% 
  column_to_rownames(var="dummy") %>% 
  as.matrix()

phy_tax <- tax_table(tax_matrix)

# Metadata import
meta_fix <- meta %>% 
  mutate(dummy=SampleID) %>%
  column_to_rownames(var="dummy")

phy_meta <- sample_data(meta_fix)

# Create phyloseq object
physeq <- phyloseq(phy_count, phy_tax, phy_meta, tree)

physeq
```

# Remove non-prokaryotic ASVs

Mitochondrial, chloroplast, protist, and unknown domain/phylum sequences.

```{r}
# Remove non-prokaryotic ASVs
physeq2 <- subset_taxa(physeq, !(Order=="Chloroplast" | Family=="Mitochondria" | Domain=="Unassigned" | 
                                   Domain=="putative Unassigned" | Domain=="putative Bacteria" | Phylum == "unclassified Bacteria"))
physeq2
```

# Sparcity filter

Keep only ASVs that occur in at least three samples.

```{r}
# Convert counts to 0 or 1 based on presence in sample
physeq_presabs <- transform_sample_counts(physeq2, function(x) ifelse(x > 0, 1, 0))
  
# Prune taxa that occur in less than three samples
physeq_presabs2 <- prune_taxa(taxa_sums(physeq_presabs) > 2, physeq_presabs) 

sparse_pass_asvs <- taxa_names(physeq_presabs2)

physeq3 <- prune_taxa(sparse_pass_asvs, physeq2)
physeq3
```

# Resequenced samples comparison

Compare sequence depth, Aquifex spike-in proportion, ASV richness.

Calculate read depth:

```{r}
# Extract count and metadata from processed phyloseq object
count_processed <- data.frame(otu_table(physeq3))

meta_processed <- data.frame(sample_data(physeq3)) %>% 
  select(everything(), -SampleID) %>% 
  rownames_to_column(var="SampleID")

# Calculate read depth per sample
read_depth <- count_processed %>%
  colSums() %>% 
  as.data.frame() %>% 
  rownames_to_column(var="SampleID") %>% 
  inner_join(meta_processed) %>% 
  rename_at('.', ~'reads')
  
# Visualize
read_depth %>%
  mutate(Replicate = as_factor(Replicate)) %>%
  #filter(Type != "pcr_control") %>%
  ggplot(aes(x=reorder(SampleID, -reads), y=reads, fill=Replicate)) +
  geom_col() +
  geom_hline(yintercept=9000) +
  labs(x="Sample", y="Reads") +
  theme_test()
```

```{r}
read_depth %>% filter(reads < 9000)
```

# Rarefy

Rarefy to 8770, lose no samples but negative controls.

```{r}
set.seed(3)

physeq_rare <- rarefy_even_depth(physeq3, sample.size=8770)

# Ag
subset_samples(physeq_rare, Soil=="Ag")
# Meadow
subset_samples(physeq_rare, Soil=="Meadow")
```

# Save preprocessed count data

Warning: Aquifex internal standard not removed.

```{r, eval=FALSE}
# Rarefied
saveRDS(physeq_rare, file="../data_intermediate/NIFA_physeq_rare.rds")

# Not rarefied
saveRDS(physeq3, file="../data_intermediate/NIFA_physeq_unrare.rds")
```

# Normalize to the internal standard

Taxon counts / Aquifex counts in each sample

```{r}
# Extract count data from rarefied phyloseq object
count_rare <- data.frame(otu_table(physeq_rare)) %>% 
  rownames_to_column(var="ASV") %>% 
  pivot_longer(!ASV, names_to="SampleID", values_to="count")

# Aquifex internal std ASVs
aqui_asvs <- filter(tax, Genus == "Aquifex")
aqui_asvs <- as.character(aqui_asvs$ASV)

# Combine int std ASVs into one ASV
count_stdsum <- count_rare %>% 
  mutate(ASV_std = if_else(ASV %in% aqui_asvs, "int_std", ASV)) %>%
  group_by(SampleID, ASV_std) %>% 
  summarize(count = sum(count)) %>% 
  ungroup()

# Divide taxa counts by internal std counts in each sample
count_norm <- count_stdsum %>% 
  pivot_wider(id_cols=SampleID, names_from = ASV_std, values_from = count) %>%
  mutate(across(-SampleID, ~ .x/int_std)) %>%
  select(-int_std) # remove Aquifex ASVs

# Merge and reformat
norm_meta <- count_norm %>% 
  pivot_longer(-SampleID, names_to="ASV", values_to="norm_abund") %>% 
  inner_join(meta_processed, by="SampleID") %>% 
  select(SampleID, Microcosm, ASV, Soil, Isotope, Replicate, Time_point, Day, norm_abund)

# Remove samples with infinite values
norm_meta <- norm_meta[!is.infinite(norm_meta$norm_abund),] %>% 
  mutate(SampleID = gsub("sa(.+)", "\\1", SampleID), # change back to numeric
         SampleID = as.numeric(SampleID))
```

# Save normalized data

```{r, eval=FALSE}
# Save normalized abundances
saveRDS(norm_meta, file="../data_intermediate/NIFA_normalized.rds")
```

```{r, eval=FALSE, echo=FALSE}
# Ben collab reformat
ben_collab <- norm_meta %>% 
  filter(Isotope=="12C") %>% 
  select(Microcosm, Replicate, Soil, Day, ASV, norm_abund)

write_tsv(ben_collab, file="../data_intermediate/ben_norm_abund.tsv")
```

