---
title: "NIFA CO2 analysis"
author: "Cassandra Wattenburger"
date: "5/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(tidyverse)

sessionInfo()

rm(list=ls())
```

# Import data

Note: Files exported from Shimadzu PostRun Analysis software as ASCII. In GCMS Postrun Analysis software highlight data files > right click > Output Items > Check Compound Quantitative Result and Qualitative Peak Table with Tab delimitation.

This script is built to run on data that has been stored in a project directory containing a "data" directory with sub-directories for each time point that .txt files with the GCMS output for each sample.

SO SLOW

```{r, eval=FALSE}
# List of directories containing data files
dir <- list.dirs(path="../data_gcms/data", full.names=TRUE, recursive=FALSE)

# Create lists of .txt files from directories
files <- list()
for (i in dir) {
  files[[i]] <- list.files(path=i, pattern="*.txt", full.names=TRUE, recursive=FALSE)
}

# Extract data
# TIC, w/z44 (12C), w/z45 (13C)
area_df <- data.frame()

for (i in names(files)) { # for each list of files (refers to a directory)
  for (x in files[[i]]) { # for each file in the selected list (directory)
    # Create sample names based on file names
    name <- str_remove(x, "../data_gcms/data/.+/") %>%
      str_remove(".txt")
    
    # Isolate time point info
    date <- str_remove(x, "../data_gcms/data/") %>%
      str_remove("/.+.txt") %>%
      str_remove(".+_")
    tp <- str_remove(x, "../data_gcms/data/") %>%
      str_remove("/.+.txt") %>%
      str_remove("_.+") %>% 
      str_remove("tp")
    
    # Isolate the area data and clean up
    dat <- read_tsv(x, skip=7, col_names=TRUE) %>%
      select("Measurement" = Name, Area) %>% # keep only the name of sample and the area measurements collected
      drop_na() %>%
      mutate(Measurement = case_when(Measurement == "TIC" ~ "TIC",
                              Measurement == "m/z 44" ~ "mz44",
                              Measurement == "m/z 45" ~ "mz45"),
             TP = tp, # adding additional metadata...
             Date = date,
             Sample = name)
    area_df <- rbind(area_df, dat) # add to data frame
  }
}
```

Save so I don't have to do that again:

```{r, eval=FALSE}
saveRDS(area_df, file="../data_intermediate/NIFA_rawgcms.rds")
```

Import:

```{r}
area_df <-readRDS("../data_intermediate/NIFA_rawgcms.rds")
```


# Reformat

```{r}
# Reformat
area_raw <- area_df %>%
  filter(Sample != "foo") %>%
  mutate(TP = as.numeric(TP),
         Type = if_else(grepl("std[0-9]+", Sample), "standard",
                        if_else(grepl("amb", Sample), "ambient", "sample")))

# Microcosm number metadata
ucosm_nums <- select(area_raw, Sample, Type) %>% 
  filter(Type != "standard" | Type != "ambient") %>% 
  mutate(Microcosm = as.numeric(Sample)) %>% 
  select(Sample, Microcosm) %>% 
  unique()

# Add metadata
meta <- read_tsv("../NIFA_metadata.tsv") 

tp_days <- meta %>% 
  select(TP=Time_point, Day) %>% 
  unique() %>% 
  filter(!is.na(TP))

soil_litter <- meta %>% 
  mutate(Litter = if_else(Isotope != "None", "litter", "no litter")) %>% 
  select(Microcosm, Soil, Litter) %>% 
  filter(Soil!="Negative")
  
area_raw <- area_raw %>% 
  left_join(tp_days) %>% 
  left_join(ucosm_nums) %>% 
  left_join(soil_litter) %>% 
  select(Sample, Microcosm, Type, Soil, Litter, TP, Date, Day, Measurement, Area)

area_raw 
```

Look at raw data:

```{r}
# all
area_raw %>% 
  ggplot(aes(x=Date, y=Area, color=Litter)) +
  geom_point(alpha=0.5) +
  facet_wrap(~Type, nrow=2) +
  geom_vline(xintercept=12, linetype=2) +
  geom_vline(xintercept=13, linetype=2) +
  #geom_text(aes(label=Microcosm), vjust="inward", hjust="inward", color="black") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45))

# Standards
area_raw %>% 
  filter(Type=="standard" & Measurement=="mz44") %>% 
  ggplot(aes(x=Date, y=Area)) +
  geom_point(alpha=0.5) +
  #geom_text(aes(label=Sample), vjust="inward", hjust="inward", color="black") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45))

# Samples
area_raw %>% 
  filter(Type=="sample") %>% 
  ggplot(aes(x=Date, y=Area, color=Measurement)) +
  geom_point(alpha=0.5) +
  facet_wrap(Litter~Measurement) +
  #geom_text(aes(label=Microcosm), vjust="inward", hjust="inward", color="black") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=6))
```

Notes:

* Stds 0-4 on day 05/29/22 are off, need to fix
* Std 4 on 06/12/22 is off
* Muc of TIC data missing, GCMS has issue detecting or calculating
* Several drop outs of mz45 data in microcosm samples, GCMS detection issue?
* One drop out of mz44 data in microcosm samples
* One abnormally heightened mz45 for no-litter sample, sample mix up?
* Missing most of sample data for 05/26/22 due to auto-injector failure

# Preprocess

Addressing abnomralities in data noted above.

Remove TIC area:

Unreliably measured.

```{r}
area_notic <- area_raw %>% 
  filter(Measurement != "TIC")
```

Standards:

Impute averages from same standards in neighboring time points.

```{r}
# Calculate imputation
imp_stds <- area_notic %>% 
  filter((Date %in% c("052822", "053122") & Sample %in% c("std0", "std1", "std2", "std3", "std4") & Measurement=="mz44") |
          (Date %in% c("061022", "061422") & Sample=="std4" & Measurement=="mz44")) %>% # neighboring stds
  mutate(group = if_else(Date %in% c("052822", "053122"), "052922", "061222")) %>% 
  group_by(Sample, group) %>% 
  summarize(Area_imp=mean(Area)) %>% 
  ungroup() %>% 
  add_column(Measurement="mz44") %>% 
  rename(Date=group)

# Impute into dataframe
area_stdimp <- area_notic %>% 
  full_join(imp_stds)

area_stdimp <- area_stdimp %>% 
  mutate(Area = if_else(is.na(Area_imp), Area, Area_imp)) %>% 
  select(Sample:Area)

# Visualize
area_stdimp %>% 
  filter(Type=="standard" & Measurement=="mz44") %>% 
  ggplot(aes(x=Date, y=Area)) +
  geom_point(alpha=0.5) +
  #geom_text(aes(label=Sample), vjust="inward", hjust="inward", color="black") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45))
```

Samples:

Remove non-detects
* Meadow 190 mz45 on 05/23/22 (13C detection issue?)
* Ag 185 mz45 on 06/12/22 (13C detection issue?)
* Ag 186 mz44 & mz45 on 05/31/22 (headspace lost?)

```{r}
# Add unique labels for easierness
area_labs <- area_stdimp %>% 
  mutate(label = paste0(Sample, Soil, Measurement, Litter, Date))

drop_labs <- c("190Meadowmz45litter052322", "185Agmz45litter061222", "186Agmz44litter053122", "186Agmz45litter053122")

area_drop <- filter(area_labs, !(label %in% drop_labs))
```

Impute lost samples on 05/26/22 via neighbor averaging
* Most samples on 05/26/22 (GCMS auto-injector failed and headspace lost)

```{r}
# Samples that were measured
area_drop %>% 
  filter(Date=="052622" & Type=="sample") %>%
  select(Sample) %>% 
  unique() %>% 
  as.list()

sams_lost <- c(187:190, 192, 193, 195)

# Calculate imputations
imp_sams <- area_drop %>% 
  filter(Microcosm %in% sams_lost & (Date=="052522" | Date=="052722")) %>% 
  group_by(Sample, Measurement) %>% 
  summarize(Area_imp = mean(Area)) %>% 
  ungroup() %>% 
  mutate(Microcosm = as.numeric(Sample),
         Litter = if_else(Microcosm %in% 185:190, "litter", "no litter"),
         Soil = if_else(Microcosm %in% c(185:186, 191:193), "Ag", "Meadow")) %>% 
  add_column(Date="052622", TP=16, Type="sample")

# Impute into dataframe
area_samimp <- area_drop %>% 
  full_join(imp_sams) %>% 
  mutate(Area = if_else(is.na(Area_imp), Area, Area_imp)) %>% 
  select(label, Sample:Area)

area_prep <- area_samimp
```

# Standards

Isolate:

```{r}
std_df <- area_prep %>%
  filter(Type=="standard") %>%
  select(label, everything())
```

Add ppm:

```{r}
# Calculate adjusted areas
std_df <- std_df %>%
  mutate(ppm = case_when(Sample == "std0" ~ 0, # add ppm
                         Sample == "std1" ~ 769,
                         Sample == "std2" ~ 1538,
                         Sample == "std3" ~ 3077,
                         Sample == "std4" ~ 7692,
                         Sample == "std5" ~ 15385,
                         Sample == "std6" ~ 30769)) 
std_df
```

Visualize:

```{r}
# mz44
std_df %>% 
  filter(Measurement=="mz44") %>% 
  ggplot(aes(x=Date, y=Area, color=Sample)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2) +
  labs(title="mz44") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45, hjust=1))

# mz45
std_df %>% 
  filter(Measurement=="mz45") %>% 
  ggplot(aes(x=Date, y=Area, color=Sample)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2) +
  labs(title="mz45") +
  theme_test() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Downward drift with time, but most pronounced in highest stds. Can they be removed?

The mz45 data is too variable at low end to be used for calibration, only mz44 will be used.

Check highest sample area vs std 5:

```{r}
sample_df <- area_prep %>% 
  filter(Type=="sample")

std_sample <- std_df %>% 
  bind_rows(sample_df)

std_sample %>% 
  filter(Measurement=="mz44") %>% 
  ggplot(aes(x=Date, y=Area, color=Type)) +
  geom_point() +
  theme_test() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Highest sample areas are lower than std 5.

Remove std 6:

```{r}
std_df <- std_df %>% 
  filter(Sample != "std6")
```

### Calibration curve

```{r}
std_cals <- tibble()
for (d in unique(std_df$Date)) {
  dat <- filter(std_df, Date==d, Measurement=="mz44")
  std_lm <- lm(Area ~ ppm, dat)
  this_row <- bind_cols(Date = unique(dat$Date), 
                       intercept = summary(std_lm)$coefficients[1,1],
                       slope = summary(std_lm)$coefficients[2,1], 
                       adjR2 = summary(std_lm)$adj.r.squared, 
                       pval = summary(std_lm)$coefficients[2,4])
  std_cals = bind_rows(std_cals, this_row)
}

std_cals
```

All look good.

# Ambients

Isolate ambient samples:

```{r}
amb_df <- area_prep %>% 
  filter(Type=="ambient") %>% 
  mutate(ppm = 0) # dummy variable for graphing
```

Visualize:

```{r}
# Against standards
bind_rows(std_df, amb_df) %>%
  filter(Measurement == "mz44") %>% 
  ggplot(aes(x=ppm, y=Area, color=Type)) +
  geom_point() +
  facet_wrap(~Day) +
  theme_test()

# alone
amb_df %>% 
  filter(Measurement=="mz44") %>% 
  ggplot(aes(x=Day, y=Area)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_line(aes(group=Date)) +
  theme_test()
```

Atmospheric samples are very low, as expected. Ambient 2 is elevated, likely due to slight carry over from std 6, also expected.

# Samples

Convert sample areas into ppm:

```{r}
sample_ppm <- sample_df %>% 
  left_join(std_cals, by="Date") %>%
  select(everything(), -adjR2, -pval) %>% 
  mutate(ppm = (Area - intercept)/slope, # convert area to ppm based on curves
         ppm = if_else(ppm < 0, 0, ppm)) # Replace negative values with 0s
```

Visualize:

```{r}
sample_ppm %>% 
  filter(Type=="sample") %>% 
  ggplot(aes(x=Day, y=ppm, color=Measurement)) +
  geom_point() +
  facet_wrap(Litter~Soil) +
  theme_test()
```

## Calculate mg CO2 & C

```{r}
# Ideal gas law: n=(P*V)/(R*temp)
# NOTE: pressure does not matter for this calulation because it's all relative to the stds
R <- 0.082057338 # universal gas constant, (L*atm)/(mol*K)
temp <- 296.48 # K, 74F in lab throughout experiment
V_m <- 0.1 # L, headspace of 100 mL serum bottle with soil, meadow
V_ag <- 0.104 # L, headspace of 100 mL serum bottle with soil, ag
mwCO2 <- 44.01 # molecular weight (g/mol) of CO2
P <- 1 # atm, assumes 1 atm in ucosms during sampling

# mol CO2 if headspace were 100% CO2
mol_ag100 <- P*V_ag/(R*temp)
mol_m100 <- P*V_m/(R*temp)

# mg CO2 if headspace were 100% CO2
mg_ag100 <- mol_ag100*mwCO2*1000
mg_m100 <- mol_m100*mwCO2*1000

# Convert ppm to mass
sample_mg <- sample_ppm %>% 
  mutate(mg_co2 = if_else(Soil=="Ag", mg_ag100*(ppm*10^-6), mg_m100*(ppm*10^-6)),
         mg_c = mg_co2*0.273) %>% 
  select(Microcosm, Soil, Litter, Day, Measurement, mg_co2, mg_c)
```

Visualize:
```{r}
sample_mg %>% 
  ggplot(aes(x=Day, y=mg_co2, color=Measurement)) +
  geom_point() +
  facet_wrap(Litter~Soil) +
  theme_test()
```

Save processed data:

```{r, eval=FALSE}
saveRDS(sample_mg, "../data_intermediate/NIFA_repiration.rds")
```

## Respiration rate

```{r}
# Time (days) elapsed per time point
time_steps <- data.frame(time_step = c(0, rep(1, 11), rep(2, 8), 3),
                         Day = sort(unique(sample_mg$Day)))

# Calculate
sample_mg <- sample_mg %>% 
  left_join(time_steps) %>% 
  filter(Day != 0) %>% 
  mutate(resp_rate_c = mg_c / time_step,
         resp_rate_co2 = mg_co2 / time_step)

# Visualize
sample_mg %>% 
  ggplot(aes(x=Day, y=resp_rate_c, color=Measurement)) +
  geom_point() +
  facet_wrap(Litter~Soil) +
  labs(title="Respiration rate (mg C)") +
  theme_test()
```

## Total respiration

```{r}
# Sum across time points
resp_total <- sample_mg %>% 
  pivot_longer(!Microcosm:Measurement, names_to="molecule", values_to="mg") %>% 
  mutate(molecule = if_else(grepl("co2", molecule), "co2", "c")) %>% 
  group_by(Microcosm, Soil, Litter, Measurement, molecule) %>% # sum over time
  summarize(mg = sum(mg)) %>% 
  ungroup() 

# Visualize
resp_total %>% 
  filter(molecule=="c") %>% 
  ggplot(aes(x=Soil, y=mg, color=Measurement)) +
  geom_point() +
  geom_vline(xintercept = 0.1) +
  facet_wrap(~Litter) +
  labs(title="Total C respiration") +
  theme_test()
```

# PIN: what is up with the low meadow oultier in the litter treatment? calculation mistake?

Statistics:

```{r}
# Litter
## mz 44
littermz44_lm <- lm(mg ~ Soil, data=filter(resp_total, Litter=="litter" & Measurement=="mz44" & molecule=="c"))
hist(resid(littermz44_lm)) # non-normal w/ or w/o log transform
plot(predict(littermz44_lm), resid(littermz44_lm)) # heteroskedastic, worse with log transform
## Switch to welch t-test
t.test(mg ~ Soil, data=filter(resp_total, Litter=="litter" & Measurement=="mz44" & molecule=="c"), var.equal=FALSE)

## mz 45
littermz45_lm <- lm(mg ~ Soil, data=filter(resp_total, Litter=="litter" & Measurement=="mz45" & molecule=="c"))
hist(resid(littermz45_lm)) # non-normal w/ or w/o log transform
plot(predict(littermz45_lm), resid(littermz44_lm)) # heteroskedastic, worse with log transform
## Switch to welch t-test
t.test(mg ~ Soil, data=filter(resp_total, Litter=="litter" & Measurement=="mz45" & molecule=="c"), var.equal=FALSE)

# No litter
## mz 44
nolittermz44_lm <- lm(mg ~ Soil, data=filter(resp_total, Litter=="no litter" & Measurement=="mz44" & molecule=="c"))
hist(resid(nolittermz44_lm)) # non-normal w/ or w/o log transform
plot(predict(nolittermz44_lm), resid(littermz44_lm)) # heteroskedastic, worse with log transform
## Switch to welch t-test
t.test(mg ~ Soil, data=filter(resp_total, Litter=="no litter" & Measurement=="mz44" & molecule=="c"), var.equal=FALSE)

## mz 45
nolittermz45_lm <- lm(mg ~ Soil, data=filter(resp_total, Litter=="no litter" & Measurement=="mz45" & molecule=="c"))
hist(resid(nolittermz45_lm)) # non-normal w/ or w/o log transform
plot(predict(nolittermz45_lm), resid(littermz44_lm)) # heteroskedastic, worse with log transform
## Switch to welch t-test
t.test(mg ~ Soil, data=filter(resp_total, Litter=="no litter" & Measurement=="mz45" & molecule=="c"), var.equal=FALSE)
```

# SCRAP

Test for autocorrelation in time series:

```{r}
ts_190M13CO2 <- sample_mg %>% 
  filter(Microcosm==190 & isotope=="labelled" & molecule=="CO2" & Soil=="Meadow") %>% # isolate Meadow 13CO2 time series
  filter(Date!="052322") %>% # remove drop out time point for autocorrelation test
  select(mg) %>% # convert into ts object type
  as.ts() 

ts_186Ag13CO2 <- sample_mg %>% 
  filter(Microcosm==186 & isotope=="labelled" & molecule=="CO2" & Soil=="Ag") %>% # isolate Meadow 13CO2 time series
  filter(Date!="052322") %>% # remove drop out time point for autocorrelation test
  select(mg) %>% # convert into ts object type
  as.ts()
  
# Autocorrelation
pacf(ts_190M13CO2)
pacf(ts_186Ag13CO2)
```

Could this be used to test for oscillation in bacterial abundance data?


