---
title: "NIFA - Co2 vs growth"
author: "Cassandra Wattenburger"
date: "4/17/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
#library(tidyverse)
library("tidyr")
library("dplyr")
library("tibble")
library("readr")
library("forcats")
library("ggplot2")

sessionInfo()

rm(list=ls())
```

# Import data

```{r}
# Respiration
resp <- readRDS("../data_intermediate/NIFA_repiration.rds")

# Normalized abundances
abund <- readRDS("../data_intermediate/NIFA_normalized.rds")

# Growth
growth <- readRDS("../data_intermediate/NIFA_grests_final_ind.rds") %>% 
  as.tibble()

# Paprica estimates
pap <- readRDS("../data_intermediate/NIFA_paprica_ests.rds")

# Metadata
meta <- read_tsv("../NIFA_metadata.tsv")
```

Reformat:

```{r}
# Only keep time series with growth estimates
abund_gr <- abund %>% 
  inner_join(growth) %>% 
  filter(!is.na(ASV) & Isotope=="12C") %>% 
  left_join(pap) %>%
  mutate(norm_abund_corr = norm_abund/n16S) %>% 
  select(ASV, Soil, Replicate, Day, norm_abund, norm_abund_corr, g, start_day, end_day, slope, yint, n16S)

abund_gr

# Add Replicate info to respiration data
ucosm_reps <- select(meta, Microcosm, Replicate)
resp <- inner_join(resp, ucosm_reps) 
```

# Total growing community abundances

```{r}
# Whole community (uncorrected)
abund %>% 
  group_by(Soil, Replicate, Day) %>% 
  summarize(norm_abund = sum(norm_abund)) %>% # sum ASV abundances
  ungroup() %>% 
  ggplot(aes(x=Day, y=norm_abund, color=Soil)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_test()

# Growing community
abund_gr %>% 
  group_by(Soil, Replicate, Day) %>% 
  summarize(norm_abund = sum(norm_abund_corr)) %>% # sum ASV abundances
  ungroup() %>% 
  ggplot(aes(x=Day, y=norm_abund, color=Soil)) +
  geom_point() +
  geom_smooth() +
  theme_test()
```

# Respiration rates

```{r}
# Time (days) elapsed per time point
time_steps <- data.frame(time_step = c(0, rep(1, 11), rep(2, 8), 3),
                         Day = sort(unique(resp$Day)))

# Calculate respiration rate mg/days
resp_rate <- resp %>% 
  left_join(time_steps) %>% 
  filter(Day != 0) %>% 
  mutate(resp_rate_c = mg_c / time_step,
         resp_rate_co2 = mg_co2 / time_step)
```

Visualize:

```{r}
resp_rate %>% 
  ggplot(aes(x=Day, y=resp_rate_c, color=Measurement)) +
  geom_point() +
  facet_wrap(Litter~Soil) +
  labs(title="Respiration rate (mg C)") +
  theme_test()

resp_rate %>% 
  filter(Litter=="litter") %>% 
  ggplot(aes(x=Day, y=resp_rate_c, color=Measurement)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~Soil) +
  labs(title="Respiration rate (mg C)") +
  theme_test()
```

# Proportions labelled CO2

```{r}
resp_total <- resp %>% 
  filter(Litter=="litter") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(mg_co2_total = sum(mg_co2),
            mg_c_total = sum(mg_c))
  

resp_prop <- resp %>% 
  filter(Litter=="litter") %>% 
  group_by(Soil, Replicate, Measurement) %>% 
  summarize(mg_co2 = sum(mg_co2),
            mg_c = sum(mg_c)) %>% # sum across all days
  ungroup() %>% 
  inner_join(resp_total) %>% 
  mutate(prop_co2 = mg_co2/mg_co2_total,
         prop_c = mg_c/mg_c_total)

resp_prop
```

Visualize:

```{r}
resp_prop %>% 
  ggplot(aes(x=Soil, y=prop_co2, color=Measurement)) +
  geom_point() +
  theme_test()
```

# Community weighted g

Based on abundance at given time point.

```{r}
# Calculate weighted g
weighted_g <- data.frame()
for (s in unique(abund_gr$Soil)) {
  for (r in unique(abund_gr[abund_gr$Soil==s,]$Replicate)) {
    for (d in unique(abund_gr[abund_gr$Soil==s & abund_gr$Replicate,]$Day)) {
      data_srd <- filter(abund_gr, Soil==s & Replicate==r & Day==d) %>%
        filter(start_day <= d & end_day >= d) # include only ASVs that grew on the given day
      weighted_mean <- weighted.mean(data_srd$g, data_srd$norm_abund_corr) # weight by corrected abundance
      this_row <- bind_cols(Soil=s, Replicate=r, Day=d, weighted_g=weighted_mean)
      weighted_g <- bind_rows(weighted_g, this_row)
    }
  }
}

weighted_g
```

Visualize:

```{r}
weighted_g %>% 
  ggplot(aes(x=Day, y=weighted_g, color=Soil)) +
  geom_point() +
  geom_smooth() +
  labs(y="Growing community weighted generation time (days)") +
  theme_test()
```

Statistics:

```{r}
# Compare soils
weightedg_lm <- lm(log(weighted_g) ~ Soil, data=weighted_g)
hist(resid(weightedg_lm)) # normal
plot(predict(weightedg_lm), resid(weightedg_lm)) # homoskedastic
summary(weightedg_lm)
```

# Respiration rate x weighted growth rate:

```{r}
# separate mz
weightedg_resprate <- inner_join(weighted_g, resp_rate) %>% 
  filter(Litter=="litter") %>% # remove samples without litter added
  select(everything(), -Litter)

weightedg_resprate

# total
weightedg_resprate_tot <- weightedg_resprate %>% 
  pivot_wider(id_cols=c(Soil, Replicate, Day, time_step, weighted_g), names_from = Measurement, values_from=mg_co2) %>% 
  drop_na() %>% 
  mutate(mgCO2 = mz44 + mz45,
         mgCO2_day = mgCO2/time_step)
weightedg_resprate_tot
```

Visualize:

```{r}
# Labelled
weightedg_resprate %>% 
  filter(Measurement=="mz45") %>%  # labelled CO2
  group_by(Soil, Day) %>%  # average across replicates
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            resp_rate_co2 = mean(resp_rate_co2, na.rm=TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(y=resp_rate_co2, x=weighted_g, color=Soil)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2) +
  theme_test() +
  labs(y="Respiration rate, mz45 (mgCO2/day)", x="w.mean g")

# Unlabelled
weightedg_resprate %>% 
  filter(Measurement=="mz44") %>%  # unlabelled CO2
  group_by(Soil, Day) %>%  # average across replicates
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            resp_rate_co2 = mean(resp_rate_co2, na.rm=TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(y=resp_rate_co2, x=weighted_g, color=Soil)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2) +
  theme_test() +
  labs(y="Respiration rate, mz44 (mgCO2/day)", x="w.mean g")

# Combined
weightedg_resprate_tot %>% 
  group_by(Soil, Day) %>% 
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            mgCO2_day = mean(mgCO2_day)) %>% 
  ungroup() %>%
  ggplot(aes(y=mgCO2_day, x=weighted_g)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2) +
  theme_test()
```

Nice figure:

```{r, eval=FALSE}
wgresp_plot <- weightedg_resprate_tot %>% 
  group_by(Soil, Day) %>% 
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            mgCO2_day = mean(mgCO2_day)) %>% 
  ungroup() %>%
  mutate(Soil = if_else(Soil=="Ag", "Agricultural", "Meadow")) %>% 
  ggplot(aes(y=mgCO2_day, x=weighted_g)) +
  geom_point(aes(shape=Soil), size=2) +
  scale_shape_manual(values=c(1,20)) +
  geom_smooth(aes(group=Soil, linetype=Soil), method="lm", color="black", alpha=0.3) +
  scale_linetype_manual(values=c("dashed", "solid")) +
  theme_test() +
  theme(axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    legend.position = "none")
wgresp_plot
```

Export:

```{r, eval=FALSE}
ggsave(wgresp_plot, file="../figures/fig_wgresp.svg", units="mm", height=70, width=80, device="svg")
```

## Statistics

Total:

```{r}
wgrr_avg_tot <- weightedg_resprate_tot %>%
  group_by(Soil, Day) %>%  # average across replicates
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            resp_rate_co2 = mean(mgCO2_day, na.rm=TRUE)) %>% 
  ungroup()

# Overall
cor.test(wgrr_avg_tot$weighted_g, wgrr_avg_tot$resp_rate_co2)

# Sep soil
cor.test(wgrr_avg_tot[wgrr_avg_tot$Soil=="Ag",]$weighted_g, wgrr_avg_tot[wgrr_avg_tot$Soil=="Ag",]$resp_rate_co2)
cor.test(wgrr_avg_tot[wgrr_avg_tot$Soil=="Meadow",]$weighted_g, wgrr_avg_tot[wgrr_avg_tot$Soil=="Meadow",]$resp_rate_co2)
```

Seperate measurements:

```{r}
# Average replicates
wgrr_avg <- weightedg_resprate %>%
  group_by(Soil, Day, Measurement) %>%  # average across replicates
  summarize(weighted_g = mean(weighted_g, na.rm=TRUE),
            resp_rate_co2 = mean(resp_rate_co2, na.rm=TRUE)) %>% 
  ungroup()

# Labelled
# Ag
wgrr_ag45_pearson <- cor.test(wgrr_avg[wgrr_avg$Measurement=="mz45" & wgrr_avg$Soil=="Ag",]$resp_rate_co2,
                         wgrr_avg[wgrr_avg$Measurement=="mz45" & wgrr_avg$Soil=="Ag",]$weighted_g, method="pearson")
wgrr_ag45_pearson
# Meadow
wgrr_mead45_pearson <- cor.test(wgrr_avg[wgrr_avg$Measurement=="mz45" & wgrr_avg$Soil=="Meadow",]$resp_rate_co2,
                         wgrr_avg[wgrr_avg$Measurement=="mz45" & wgrr_avg$Soil=="Meadow",]$weighted_g, method="pearson")
wgrr_mead45_pearson

# Unlabelled
# Ag
wgrr_ag44_pearson <- cor.test(wgrr_avg[wgrr_avg$Measurement=="mz44" & wgrr_avg$Soil=="Ag",]$resp_rate_co2,
                         wgrr_avg[wgrr_avg$Measurement=="mz44" & wgrr_avg$Soil=="Ag",]$weighted_g, method="pearson")
wgrr_ag44_pearson
# Meadow
wgrr_mead44_pearson <- cor.test(wgrr_avg[wgrr_avg$Measurement=="mz44" & wgrr_avg$Soil=="Meadow",]$resp_rate_co2,
                         wgrr_avg[wgrr_avg$Measurement=="mz44" & wgrr_avg$Soil=="Meadow",]$weighted_g, method="pearson")
wgrr_mead44_pearson

# Multiple test correction
p.adjust(p=c(wgrr_ag45_pearson$p.value, wgrr_mead45_pearson$p.value, wgrr_ag44_pearson$p.value, wgrr_mead44_pearson$p.value),
           method="holm", n=4)
```

# Growth efficiency

Proxy for CUE using generation time, period of time, normalized abundance, and respiration rate.

N0 = normalized abundance of taxon at start of time interval (not log transformed)
g = generation time of taxon (days)
t = days in time frame
r = mg CO2/day, respiration rate

n = t/g, number of generations
Nt = N0*2^n, biomass at end of time frame
deltaN = Nt - N0, change in biomass
r*t = the mg CO2 produced in that time frame

growth efficiency = ΣdeltaN/(r*t), ie sum of gained abundance divided by mg CO2 produced.

## Impute abundances

Based on linear growth estimate, using imputed data to prevent bias from drop outs.

```{r}
# Respiration headspace sampling time points
resp_tps <- unique(resp$Day) %>% 
  data.frame(day=.) %>% 
  arrange(day) %>% 
  add_column(tp=c(1:nrow(.)))

# Impute abundance
abund_impute <- abund_gr %>% 
  mutate(growing = if_else(Day >= start_day & Day <= end_day, "yes", "no"),  # is it growing during given time interval?
         impute_abund = if_else(growing=="yes", slope*Day+yint, NA), # impute using linear model, creates ln imputed abund
         impute_abund_corr = exp(impute_abund)/n16S, # un-ln and correct for n16S
         label = paste0(Soil, "_", Replicate, "_", ASV)) # add label

abund_impute

# Calculate change in imputed abundance for each headspace time interval
resp_intends <- resp_tps[2:nrow(resp_tps),]
chabund_df <- data.frame()
for (l in unique(abund_impute$label)) { # for each time series
  #print(l)
  data_l <- filter(abund_impute, label==l) %>%
    arrange(Day) %>% 
    filter(growing=="yes") %>% 
    select(label, day=Day, impute_abund_corr)
  
  for (d in resp_intends$day) { # for each headspace sampling day (not including day 0)
    start_int <- resp_tps[resp_tps[resp_tps$day==d,]$tp-1,]$day
    #print(paste0("Start day " , start_int))
    #print(paste0("End day ", d))
    
    # Growing on start and end days of time interval
    if (start_int %in% data_l$day & d %in% data_l$day) {
      #print("Growing on start and end days of time interval")
      abund1 <- NULL; abund2 <- NULL; chabund <- NULL
      abund1 <- data_l %>% 
        filter(day==start_int) %>% 
        select(impute_abund_corr) %>% 
        as.numeric()
      abund2 <- data_l %>% 
        filter(day==d) %>% 
        select(impute_abund_corr) %>% 
        as.numeric()
      chabund <- abund2-abund1
      this_row <- data.frame(label=l, start=start_int, end=d, chabund)
      chabund_df <- bind_rows(chabund_df, this_row)
      #print(chabund)
    }
    # Not growing on start or end
    else {
      # Growing on start day but not end day of time interval
      ## this means either ended growth on start day or stopped growth on an intermediate tp
      if (start_int %in% data_l$day) {
        #print("Growing on start day but not end day of time interval")
        abund1 <- NULL; abund2 <- NULL; chabund <- NULL
        abund1 <- data_l %>% 
          filter(day==start_int) %>% 
          select(impute_abund_corr) %>% 
          as.numeric()
        abund2 <- data_l %>% 
          summarize(max(impute_abund_corr)) %>% # growth ended before end of time interval, so largest abundance is terminal
          as.numeric()
        chabund <- abund2 - abund1
        this_row <- data.frame(label=l, start=start_int, end=d, chabund)
        chabund_df <- bind_rows(chabund_df, this_row)
        #print(chabund)
      }
      # Growing on end day of time interval but not start day
      ## this means growth likely started at the end of the time interval or on an intermediate time point
      else if (d %in% data_l$day) {
        #print("Growing on end day of time interval but not start day")
        abund1 <- NULL; abund2 <- NULL; chabund <- NULL
        abund1 <- data_l %>% 
          summarize(min(impute_abund_corr)) %>%  # growth started after start of time interval so minimum abundance is start
          as.numeric()
        abund2 <- data_l %>% 
          filter(day==d) %>% 
          select(impute_abund_corr) %>% 
          as.numeric()
        chabund <- abund2 - abund1
        this_row <- data.frame(label=l, start=start_int, end=d, chabund)
        chabund_df <- bind_rows(chabund_df, this_row)
        #print(chabund)
      }
      # Didn't grow at all during time interval
      else {
        #print("Didn't grow at all during time interval")
        abund1 <- NULL; abund2 <- NULL; chabund <- NULL
        next
      }
    }
  }
}

meta_impute <- abund_impute %>% 
  select(label, Soil, Replicate, ASV) %>% 
  unique()

# Sum across ASVs
chabund_sum <- chabund_df %>%
  left_join(meta_impute) %>% 
  group_by(Soil, Replicate, start, end) %>% 
  summarize(chabund = sum(chabund)) %>% 
  ungroup()

chabund_sum
```

## Calculate growth eff.

```{r}
# Add replicate numbers to resp
meta2 <- meta %>% 
  select(Microcosm, Soil, Replicate) %>% 
  unique()

resp2 <- left_join(resp, meta2)

# Calculate
greff_df <- chabund_sum %>% 
  select(Soil, Replicate, Day=end, chabund) %>% 
  inner_join(resp2) %>% 
  unique() %>% 
  filter(Litter=="litter") %>% 
  select(Soil, Replicate, Day, chabund, measurement=Measurement, mg_co2) %>% 
  group_by(Soil, Replicate, Day, chabund) %>% # sum labelled and unlabelled respiration
  summarize(mg_co2 = sum(mg_co2)) %>% 
  ungroup() %>% 
  mutate(greff = chabund/mg_co2)

greff_df
```

Visualize:

```{r}
greff_df %>% 
  ggplot(aes(x=Day, y=log(greff), linetype=Soil)) +
  geom_point(aes(shape=Soil), size=2) +
  scale_shape_manual(values=c(1,16)) +
  geom_smooth(color="black") +
  scale_linetype_manual(values=c(2,1)) +
  theme_test()
```

```{r}
greff_plot_legend <- greff_df %>% 
  mutate(Soil=if_else(Soil=="Ag", "Agricultural", "Meadow")) %>% 
  ggplot(aes(x=Day, y=log(greff), linetype=Soil)) +
  geom_point(aes(shape=Soil), size=2) +
  scale_shape_manual(values=c(1,16)) +
  geom_smooth(color="black") +
  scale_linetype_manual(values=c(2,1)) +
  theme_test() +
  theme(axis.title = element_text(size=8),
      axis.text = element_text(size=8),
      legend.title = element_text(size=8),
      legend.text = element_text(size=8))
greff_plot_legend

greff_plot_nolegend <- greff_df %>% 
  mutate(Soil=if_else(Soil=="Ag", "Agricultural", "Meadow")) %>% 
  ggplot(aes(x=Day, y=log(greff), linetype=Soil)) +
  geom_point(aes(shape=Soil), size=2) +
  scale_shape_manual(values=c(1,16)) +
  geom_smooth(color="black") +
  scale_linetype_manual(values=c(2,1)) +
  theme_test() +
  theme(axis.title = element_text(size=8),
      axis.text = element_text(size=8),
      legend.position = "none")
greff_plot_nolegend
```

Export:

```{r, eval=FALSE}
ggsave(greff_plot_legend, file="../figures/fig_greff_legend.svg", units="mm", height=70, width=85, device="svg")

ggsave(greff_plot_nolegend, file="../figures/fig_greff_nolegend.svg", units="mm", height=70, width=85, device="svg")
```

## Statistics

```{r}
greff_lm <- lm(greff ~ Soil*Day, data=greff_df)
hist(resid(greff_lm)) # normal
plot(predict(greff_lm), resid(greff_lm)) # heteroskedastic
# Switch to welch
oneway.test(greff ~ Soil, data=greff_df, var.equal=FALSE)
oneway.test(greff ~ Day, data=greff_df, var.equal=FALSE)
```

Export growth efficiency calculations:

```{r, eval=fALSE}
saveRDS(greff_df, file="../data_intermediate/NIFA_greff.rds")
```

# Weighted g vs greff

```{r}
greff_df %>% 
  inner_join(weighted_g) %>% 
  group_by(Soil, Day) %>% 
  summarize(greff = mean(greff),
            weighted_g = mean(weighted_g)) %>%
  ungroup() %>% 
  ggplot(aes(x=weighted_g, y=log(greff))) +
  geom_point(aes(shape=Soil)) +
  scale_shape_manual(values=c(1,16)) +
  geom_smooth(method="lm", color="black", linetype=2) +
  theme_test()
```

Nice figure:

```{r, eval=FALSE}
wggreff_plot <- greff_df %>% 
  inner_join(weighted_g) %>% 
  group_by(Soil, Day) %>% 
  summarize(greff = mean(greff),
            weighted_g = mean(weighted_g)) %>%
  ungroup() %>% 
  mutate(Soil=if_else(Soil=="Ag", "Agricultural", "Meadow")) %>% 
  ggplot(aes(x=weighted_g, y=log(greff))) +
  geom_point(aes(shape=Soil), size=2) +
  scale_shape_manual(values=c(1,16)) +
  geom_smooth(aes(group=Soil, linetype=Soil), method="lm", color="black", alpha=0.3) +
  scale_linetype_manual(values=c("dashed", "solid")) +
  theme_test() +
  theme(axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    legend.position = "none")
wggreff_plot
```

```{r, eval=FALSE}
ggsave(wggreff_plot, file="../figures/fig_wggreff.svg", units="mm", height=70, width=80, device="svg")
```

#### Statistics

```{r}
greff_avg <- greff_df %>% 
  inner_join(weighted_g) %>% 
  group_by(Soil, Day) %>% 
  summarize(greff = mean(greff),
            weighted_g = mean(weighted_g)) %>%
  ungroup()

# Overall
cor.test(greff_avg$weighted_g, log(greff_avg$greff))

# Sep soils
cor.test(greff_avg[greff_avg$Soil=="Ag",]$weighted_g, log(greff_avg[greff_avg$Soil=="Ag",]$greff))
cor.test(greff_avg[greff_avg$Soil=="Meadow",]$weighted_g, log(greff_avg[greff_avg$Soil=="Meadow",]$greff))
```

# Greff vs respiration rate

```{r}
greff_df %>% 
  inner_join(weightedg_resprate_tot) %>% 
  group_by(Soil, Day) %>% 
  summarize(greff = mean(greff),
            mgCO2_day = mean(mgCO2_day)) %>% 
  ungroup() %>%
  ggplot(aes(x=mgCO2_day, y=log(greff))) +
  geom_point() +
  theme_test()
```

