---
title: "NIFA soil treatment analysis"
author: "Cassandra Wattenburger"
date: "2/17/2023"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
```

```{r}
# Clear working directory, load in packages, generate package info
rm(list=ls())

library("tidyverse")
library("phyloseq")
library("vegan")
library("ape")

sessionInfo()
```

```{r}
# ggplot theme
# See: https://benjaminlouis-stat.fr/en/blog/2020-05-21-astuces-ggplot-rmarkdown/
theme_clean <- function(base_size = 14) {
  theme_test(base_size = base_size) %+replace%
    theme(axis.title = element_text(size=12),
          axis.text = element_text(size=10, color="black"),
          legend.text = element_text(size=10),
          legend.title = element_text(size=10),
          panel.border = element_rect(color="black"))
}
```

# Import data

```{r}
# Growth estimates
growth <- readRDS("../data_intermediate/NIFA_grests_final_ind.rds")

# Death estimates
death <- readRDS("../data_intermediate/NIFA_dests_final_ind.rds")

# Paprica estimates
pap <- readRDS("../data_intermediate/NIFA_paprica_ests.rds")

# Prepped normalized abundnace data
abund <- readRDS("../data_intermediate/NIFA_norm_prepped.rds")

# Taxonomy
tax <- read_tsv("../data_amplicon/NIFA2/final/NIFA2.taxonomy-final.tsv")

# metadata
meta <- read_tsv("../NIFA_metadata.tsv")
meta_sa <- meta %>% 
  mutate(SampleID_seq = paste0("sa", SampleID)) # match with phyloseq's coerced sample names

# Phyloseq, rarefied with no int std
physeq <- readRDS("../data_intermediate/NIFA_physeq_rare_nostd.rds")
```

Reformat data: 

```{r}
# Add paprica estimates
growth2 <- growth %>% 
  left_join(pap)

death2 <- death %>% 
  left_join(pap)

# Combine growth and death estimates
growth_death <- death2 %>%
  rename(g="h") %>% 
  bind_rows(growth2) %>% 
  mutate(type = if_else(slope < 0, "death", "growth"),
         start_abund_corr = start_abund/n16S,
         end_abund_corr = end_abund/n16S,
         change_abund_corr = change_abund/n16S) %>% 
  select(label:end_day, start_abund_corr, end_abund_corr, change_abund_corr, n16S:type)
```

# Diversity overview

## Alpha diversity

After preprocessing, sparcity filtering, and rarefication.

```{r}
# Calculate richness per sample
asv_count <- data.frame(otu_table(physeq))
richness <- asv_count %>% 
  mutate(across(everything(), ~if_else(.x > 0, 1, 0))) %>% # presence/absence
  colSums() %>% 
  data.frame() %>% 
  rename(richness=".") %>% 
  rownames_to_column(var="SampleID_seq") %>% 
  left_join(meta_sa) %>% # add metadata
  select(SampleID, Microcosm, Soil, Replicate, Day, Isotope, richness)

# Summarize for each soil
richness %>% 
  filter(Isotope=="12C") %>% 
  group_by(Soil, Isotope) %>% 
  summarize(richness_mean = mean(richness),
            richness_sd = sd(richness))

# Summarize by soil and day
richness %>% 
  filter(Isotope=="12C") %>% 
  group_by(Soil, Day, Isotope) %>% 
  summarize(richness_mean = mean(richness),
            richness_sd = sd(richness))
```

Visualize:

```{r}
richness %>% 
  filter(Isotope=="12C") %>% 
  ggplot(aes(x=Day, y=richness, color=Soil)) +
  geom_point() +
  geom_smooth(method="lm", linetype=2, lpah=0.5) +
  theme_test()
```

### Statistics

```{r}
# Soils
rich_lm <- lm(log(richness)+Day+log(richness)*Day ~ Soil, data=richness)
hist(resid(rich_lm)) # non-normal
plot(predict(rich_lm), resid(rich_lm)) # homoskedastic with log transform
# Non-parametric tests
kruskal.test(log(richness) ~ Soil, data=richness)
kruskal.test(richness ~ Day, data=richness)
```

Nice figure:

```{r, eval=FALSE}
rich_plot <- richness %>% 
  mutate(Soil = if_else(Soil=="Ag", "Agricultural", "Meadow")) %>% 
  filter(Isotope=="12C") %>% 
  ggplot(aes(x=Day, y=richness, shape=Soil)) +
  scale_shape_manual(values=c(1,16)) +
  geom_point(size=2) +
  theme_test() +
  theme(axis.title = element_text(size=14),
    axis.text = element_text(size=12),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12))
rich_plot
```

Export:

```{r, eval=FALSE}
ggsave(rich_plot, file="../figures/suppfig_richness.svg", units="mm", height=90, width=180, device="svg")
```

Total unique ASVs with growth/death estimates:

```{r}
# Growth
# Ag
growth_death %>% 
  filter(type=="growth" & Soil=="Ag") %>% 
  select(ASV) %>% 
  unique() %>% 
  nrow()

# Meadow
growth_death %>% 
  filter(type=="growth" & Soil=="Meadow") %>% 
  select(ASV) %>% 
  unique() %>% 
  nrow()

# Death
# Ag
growth_death %>% 
  filter(type=="death" & Soil=="Ag") %>% 
  select(ASV) %>% 
  unique() %>% 
  nrow()

# Meadow
growth_death %>% 
  filter(type=="death" & Soil=="Meadow") %>% 
  select(ASV) %>% 
  unique() %>% 
  nrow()
```

Phyla with most estimated ASVs?

```{r}
# Growth

# Total estimates per phylum per soil
growth_phytots <- growth_death %>% 
  inner_join(tax) %>% 
  filter(type=="growth") %>% 
  select(Soil, Phylum, ASV) %>% 
  unique() %>% 
  group_by(Soil, Phylum) %>% 
  summarize(phy_total = n()) %>% 
  ungroup()

# Total estimates per soil
growth_soiltots <- growth_phytots %>% 
  group_by(Soil) %>%
  summarise(soil_total = sum(phy_total)) %>% 
  ungroup()
  
# Relative proportion of phylum estimates
growth_soilphytots <- growth_phytots %>% 
  full_join(growth_soiltots) %>% 
  mutate(relabund = phy_total/soil_total)

# Visualize
palette_bw2 <- c("white", "darkgrey")

growth_soilphytots %>% 
  ggplot(aes(x=reorder(Phylum, -relabund), y=relabund, fill=Soil)) +
  geom_col(color="black", position = "dodge") +
  labs(y="Relative proportion", title="Proportion of growing ASVs from each phyla per soil") +
  scale_fill_manual(values=palette_bw2) +
  theme_test() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=10),
        panel.border = element_rect(color="black")
        )

# Death

# Total estimates per phylum per soil
growth_phytots <- growth_death %>% 
  inner_join(tax) %>% 
  filter(type=="death") %>% 
  select(Soil, Phylum, ASV) %>% 
  unique() %>% 
  group_by(Soil, Phylum) %>% 
  summarize(phy_total = n()) %>% 
  ungroup()

# Total estimates per soil
growth_soiltots <- growth_phytots %>% 
  group_by(Soil) %>%
  summarise(soil_total = sum(phy_total)) %>% 
  ungroup()
  
# Relative proportion of phylum estimates
growth_soilphytots <- growth_phytots %>% 
  full_join(growth_soiltots) %>% 
  mutate(relabund = phy_total/soil_total)

# Visualize
growth_soilphytots %>% 
  ggplot(aes(x=reorder(Phylum, -relabund), y=relabund, fill=Soil)) +
  geom_col(color="black", position = "dodge") +
  labs(y="Relative proportion", title="Proportion of dying ASVs from each phyla per soil") +
  scale_fill_manual(values=palette_bw2) +
  theme_test() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=10),
        panel.border = element_rect(color="black"))
```

## Beta diversity

Overall:

```{r}
set.seed(2023)

# Make tree binary (fixes this issue: https://github.com/joey711/phyloseq/issues/936)
tree <- phy_tree(physeq)
is.rooted(tree)
tree_binary <- multi2di(tree)

physeq2 <- phyloseq(sample_data(physeq), otu_table(physeq), tax_table(physeq), tree_binary)
physeq2 <- subset_samples(physeq2, Isotope=="12C") # remove 13C samples

# Distance, weighted unifrac
wuni_dist <- distance(physeq2, method="wunifrac")

# Distance, unweighted unifrac
uni_dist <- distance(physeq2, method="uunifrac")

# Ordinate, NMDS
nmds_wuni <- ordinate(physeq2, method="NMDS", distance=wuni_dist) 
nmds_uni <- ordinate(physeq2, method="NMDS", distance=uni_dist)
```

Visualize:

```{r}
# Visualize
plot_ordination(physeq2, nmds_wuni, color="Day", shape="Soil") +
  geom_point(size=2) +
  labs(title="Weighted unifrac") +
  theme_test() 

plot_ordination(physeq2, nmds_uni, color="Day", shape="Soil") +
  geom_point(size=2) +
  labs(title="Unweighted unifrac") +
  theme_test() 
```

Nice figure:

```{r, eval=FALSE}
wuni_plot <- plot_ordination(physeq2, nmds_wuni, color="Day", shape="Soil") +
  geom_point(size=2) +
  scale_shape_manual(values=c(1, 16)) +
  theme_test() +
  theme(axis.title = element_text(size=14),
    axis.text = element_text(size=12),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12))
wuni_plot
```

Export:

```{r, eval=FALSE}
ggsave(wuni_plot, file="../figures/suppfig_betadiv.svg", units="mm", width=180, height=90, device="svg")
```

### Statistics

```{r}
adonis(wuni_dist~Soil+Day+Soil*Day, data=data.frame(sample_data(physeq2)), permutations=999)
```

# Community-level

Growth summary:

```{r}
# Growth

# Average number of estimates
growth_death %>%
  filter(type=="growth") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(ests=n()) %>% 
  ungroup() %>%
  group_by(Soil) %>% 
  summarize(ests_mean = mean(ests),
            ests_sd = sd(ests))

# Average generation time
growth_death %>% 
  filter(type=="growth") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(g = mean(g)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(g_mean = mean(g),
            g_sd = sd(g))

# Lag time
growth_death %>% 
  filter(type=="growth") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(lag = mean(start_day)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(lag_mean = mean(lag),
            lag_sd = sd(lag))

# change in abundance
growth_death %>% 
  filter(type=="growth") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(chabund = mean(change_abund_corr, na.rm=TRUE)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(chabund_mean = mean(chabund),
            chabund_sd = sd(chabund))
```

Death summary:

```{r}
# Death

# Average number of estimates
growth_death %>%
  filter(type=="death") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(ests=n()) %>% 
  ungroup() %>%
  group_by(Soil) %>% 
  summarize(ests_mean = mean(ests),
            ests_sd = sd(ests))

# Average generation time
growth_death %>% 
  filter(type=="death") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(g = mean(g)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(g_mean = mean(g),
            g_sd = sd(g))

# Lag time
growth_death %>% 
  filter(type=="death") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(lag = mean(start_day)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(lag_mean = mean(lag),
            lag_sd = sd(lag))

# change in abundance
growth_death %>% 
  filter(type=="death") %>% 
  group_by(Soil, Replicate) %>% 
  summarize(chabund = mean(change_abund_corr, na.rm=TRUE)) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  summarise(chabund_mean = mean(chabund),
            chabund_sd = sd(chabund))
```

Visualize:

```{r}
# Average across ASVs per replicate
growth_death_com <- growth_death %>% 
  group_by(Soil, Replicate, type) %>% 
  summarize(k = mean(k),
            g = mean(g),
            start_day = mean(start_day, na.rm=TRUE),
            end_day = mean(end_day, na.rm=TRUE),
            start_abund_corr = mean(start_abund_corr, na.rm=TRUE),
            end_abund_corr = mean(end_abund_corr, na.rm=TRUE),
            change_abund_corr = mean(change_abund_corr, na.rm=TRUE),
            n16S = mean(n16S, na.rm=TRUE),
            genome_size = mean(genome_size, na.rm=TRUE),
            npaths_actual = mean(npaths_actual, na.rm=TRUE),
            nec_actual = mean(nec_actual, na.rm=TRUE)) %>% 
  ungroup()

# Visualize
growth_death_com %>% 
  pivot_longer(cols = -c(Soil, Replicate, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=abs(value), color=Soil)) +
  geom_jitter(width=0.1) +
  facet_wrap(~variable, scales="free") +
  theme_test()

growth_death_com %>% 
  pivot_longer(cols = -c(Soil, Replicate, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=log(abs(value)), color=Soil)) +
  geom_jitter(width=0.1) +
  facet_wrap(~variable, scales="free") +
  theme_test()
```

Figure:

```{r}
# Figure
palette_bw2 <- c("black", "white")
growth_death_com %>% 
  select(Soil, Replicate, type, g, start_day, start_abund=start_abund_corr, end_abund=end_abund_corr, change_abund=change_abund_corr, n16S) %>%
  pivot_longer(cols = -c(Soil, Replicate, type), names_to = "variable", values_to = "value") %>%
  mutate(variable = factor(variable, levels=c("start_day", "g", "n16S", "start_abund", "end_abund", "change_abund"))) %>%
  ggplot(aes(x=type, y=abs(value), shape=Soil)) +
  geom_jitter(width=0.1, size=2) +
  facet_wrap(~variable, scales="free") +
  #scale_color_manual(values=palette_bw2) +
  scale_shape_manual(values=c(1,16)) +
  theme_test() +
  theme(axis.title.x = element_blank(),
      axis.text = element_text(size=10, color="black"),
      axis.title = element_text(size=12),
      legend.text = element_text(size=10),
      legend.title = element_text(size=10),
      panel.border = element_rect(color="black"),
      strip.text = element_text(size = 10),
      strip.background = element_blank())
```

## Statistics

Growth:

```{r}
# generation time
g_gr_lm <- lm(log(g) ~ Soil, data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(g_gr_lm)) # ????
plot(predict(g_gr_lm), resid(g_gr_lm)) # homoskedastic
anova(g_gr_lm)

# start day
lag_gr_lm <- lm(start_day ~ Soil, data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(lag_gr_lm)) # non-normal
plot(predict(lag_gr_lm), resid(g_gr_lm)) # homoskedastic
# Use non-parametric
lag_gr_welch <- t.test(start_day ~ Soil, data=growth_death_com[growth_death_com$type=="growth",], var.equal=FALSE)

# n16S
n16S_gr_lm <- lm(n16S ~ Soil, data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(n16S_gr_lm)) # ????
plot(predict(n16S_gr_lm), resid(g_gr_lm)) # homoskedastic
anova(n16S_gr_lm)

# Change abund
chabund_gr_lm <- lm(change_abund_corr ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(chabund_gr_lm))
plot(predict(chabund_gr_lm), resid(chabund_gr_lm))
summary(chabund_gr_lm)

# Start abund
startabund_gr_lm <- lm(log(start_abund_corr) ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(startabund_gr_lm))
plot(predict(startabund_gr_lm), resid(startabund_gr_lm))
summary(startabund_gr_lm)

# End abund
endabund_gr_lm <- lm(log(end_abund_corr) ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="growth",])
hist(resid(endabund_gr_lm))
plot(predict(endabund_gr_lm), resid(endabund_gr_lm))
summary(endabund_gr_lm)

# Multiple test correction
growth_com_padj <- p.adjust(c(summary(g_gr_lm)$coefficients[2,4], lag_gr_welch$p.value, 
                              summary(n16S_gr_lm)$coefficients[2,4], summary(chabund_gr_lm)$coefficients[2,4], 
                              summary(startabund_gr_lm)$coefficients[2,4], summary(endabund_gr_lm)$coefficients[2,4]), 
                            method = "holm", n=6)

growth_com_padj
```

Death:

```{r}
# generation time
g_de_lm <- lm(log(g) ~ Soil, data=growth_death_com[growth_death_com$type=="death",])
hist(resid(g_de_lm)) # ????
plot(predict(g_de_lm), resid(g_de_lm)) # homoskedastic
anova(g_de_lm)

# start day
lag_de_lm <- lm(start_day ~ Soil, data=growth_death_com[growth_death_com$type=="death",])
hist(resid(lag_de_lm)) # ????
plot(predict(lag_de_lm), resid(g_de_lm)) # homoskedastic
anova(lag_de_lm)

# n16S
n16S_de_lm <- lm(log(n16S) ~ Soil, data=growth_death_com[growth_death_com$type=="death",])
hist(resid(n16S_de_lm)) # non-normal
plot(predict(n16S_de_lm), resid(n16S_de_lm)) # heteroskedastic
# Use welch t.test to deal with heteroskedasticity
n16S_de_kw <- t.test(n16S ~ Soil, data=growth_death_com[growth_death_com$type=="death",], var.equal=FALSE)

# Change abund
chabund_de_lm <- lm(log(abs(change_abund_corr)) ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="death",])
hist(resid(chabund_de_lm)) # non-normal even with log transform
plot(predict(chabund_de_lm), resid(chabund_de_lm)) # homoskedastic
# Use non-parametric test
chabund_de_kw <- kruskal.test(change_abund_corr ~ Soil, data=growth_death_com[growth_death_com$type=="death",])

# Start abund
startabund_de_lm <- lm(log(start_abund_corr) ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="death",])
hist(resid(startabund_de_lm)) # non-normal
plot(predict(startabund_de_lm), resid(startabund_de_lm)) # homoskedastic
# Use non-parametric test
startabund_de_kw <- kruskal.test(start_abund_corr ~ Soil, data=growth_death_com[growth_death_com$type=="death",])

# End abund
endabund_de_lm <- lm(log(end_abund_corr) ~ Soil, 
                 data=growth_death_com[growth_death_com$type=="death",])
hist(resid(endabund_de_lm)) # normal
plot(predict(endabund_de_lm), resid(endabund_de_lm)) # heteroskedastic
# Use welch t-test
endabund_de_welch <- t.test(end_abund_corr ~ Soil, data=growth_death_com[growth_death_com$type=="death",], equal.var=FALSE)

# Multiple test correction
death_com_padj <- p.adjust(c(summary(g_de_lm)$coefficients[2,4], summary(lag_de_lm)$coefficients[2,4],
                             n16S_de_kw$p.value, chabund_de_kw$p.value, startabund_de_kw$p.value, endabund_de_welch$p.value),
                           method = "holm", n=6)

death_com_padj
```

# ASV-level

## Metric vs soil

```{r}
# Average across replicates per ASV
growth_death_asv <- growth_death %>% 
  group_by(Soil, ASV, type) %>% 
  summarize(k = mean(k),
            g = mean(g),
            start_day = mean(start_day, na.rm=TRUE),
            end_day = mean(end_day, na.rm=TRUE),
            start_abund_corr = mean(start_abund_corr, na.rm=TRUE),
            end_abund_corr = mean(end_abund_corr, na.rm=TRUE),
            change_abund_corr = mean(change_abund_corr, na.rm=TRUE),
            n16S = mean(n16S, na.rm=TRUE),
            genome_size = mean(genome_size, na.rm=TRUE),
            npaths_actual = mean(npaths_actual, na.rm=TRUE),
            nec_actual = mean(nec_actual, na.rm=TRUE)) %>% 
  ungroup()
```

Visualize:

```{r}
# Visualize

# Growth traits
growth_death_asv %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=abs(value), color=Soil)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free") +
  theme_test()

growth_death_asv %>%
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=log(abs(value)), color=Soil)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free") +
  labs(title="Growth - log") +
  theme_test()

# Death traits
growth_death_asv %>% 
  filter(type=="death") %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=abs(value), color=Soil)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free") +
  labs(title="Growth") +
  theme_test()

growth_death_asv %>% 
  filter(type=="death") %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  ggplot(aes(x=type, y=log(abs(value)), color=Soil)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free") +
  labs(title="Growth - log") +
  theme_test()

# Growth trait density
growth_death_asv %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  filter(type=="growth") %>% 
  ggplot(aes(abs(value), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  labs(title="Growth") +
  theme_test()

growth_death_asv %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  filter(type=="growth") %>% 
  ggplot(aes(log(abs(value)), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  labs(title="Growth - log") +
  theme_test()

# Death trait density
growth_death_asv %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  filter(type=="death") %>% 
  ggplot(aes(abs(value), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  labs(title="Death") +
  theme_test()

growth_death_asv %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  filter(type=="death") %>% 
  ggplot(aes(log(abs(value)), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  labs(title="Death - log") +
  theme_test()
```

Figure:

```{r}
# Growth
palette_bw2 <- c("darkgrey", "black")
growth_death_asv %>% 
  filter(type=="growth") %>% 
  select(Soil:type, g, start_day, start_abund=start_abund_corr, end_abund=end_abund_corr, change_abund=change_abund_corr, n16S) %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels=c("start_day", "g", "n16S", "start_abund", "end_abund", "change_abund"))) %>%
  ggplot(aes(log(abs(value)), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  scale_color_manual(values=palette_bw2) +
  labs(title="Growth") +
  theme_test() +
  theme(axis.text = element_text(size=10, color="black"),
      axis.title = element_text(size=12),
      legend.text = element_text(size=10),
      legend.title = element_text(size=10),
      panel.border = element_rect(color="black"),
      strip.text = element_text(size = 10),
      strip.background = element_blank())

# Death
palette_bw2 <- c("darkgrey", "black")
growth_death_asv %>% 
  filter(type=="death") %>% 
  select(Soil:type, g, start_day, start_abund=start_abund_corr, end_abund=end_abund_corr, change_abund=change_abund_corr, n16S) %>% 
  pivot_longer(cols = -c(Soil, ASV, type), names_to = "variable", values_to = "value") %>% 
  mutate(variable = factor(variable, levels=c("start_day", "g", "n16S", "start_abund", "end_abund", "change_abund"))) %>%
  ggplot(aes(log(abs(value)), color=Soil)) +
  geom_density() +
  facet_wrap(~variable, scales="free") +
  scale_color_manual(values=palette_bw2) +
  labs(title="Death") +
  theme_test() +
  theme(axis.text = element_text(size=10, color="black"),
      axis.title = element_text(size=12),
      legend.text = element_text(size=10),
      legend.title = element_text(size=10),
      panel.border = element_rect(color="black"),
      strip.text = element_text(size = 10),
      strip.background = element_blank())
```

### Statistics

Growth ASV-level:

```{r}
# generation time
g_gr_lm <- lm(log(g) ~ Soil, data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(g_gr_lm)) # normal
plot(predict(g_gr_lm), resid(g_gr_lm)) # homoskedastic
summary(g_gr_lm)

# start day
lag_gr_lm <- lm(start_day ~ Soil, data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(lag_gr_lm)) # non-normal
plot(predict(lag_gr_lm), resid(g_gr_lm)) # homoskedastic
# Use non-parametric test
lag_gr_kw <- kruskal.test(start_day ~ Soil, data=growth_death_asv[growth_death_asv$type=="growth",])

# n16S
n16S_gr_lm <- lm(n16S ~ Soil, data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(n16S_gr_lm)) # non-normal
plot(predict(n16S_gr_lm), resid(g_gr_lm)) # homoskedastic
# Use non-parametric test
n16S_gr_kw <- kruskal.test(n16S ~ Soil, data=growth_death_asv[growth_death_asv$type=="growth",])

# Change abund
chabund_gr_lm <- lm(log(change_abund_corr) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(chabund_gr_lm)) # normal
plot(predict(chabund_gr_lm), resid(chabund_gr_lm)) # homoskedastic w/ or w/o log transform
summary(chabund_gr_lm)

# Start abund
startabund_gr_lm <- lm(log(start_abund_corr) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(startabund_gr_lm)) # normal
plot(predict(startabund_gr_lm), resid(startabund_gr_lm)) # homoskedastic
summary(startabund_gr_lm)

# End abund
endabund_gr_lm <- lm(log(end_abund_corr) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="growth",])
hist(resid(endabund_gr_lm)) # normal
plot(predict(endabund_gr_lm), resid(endabund_gr_lm)) # homoskedastic
summary(endabund_gr_lm)

# Multiple test correction
growth_asv_padj <- p.adjust(c(summary(g_gr_lm)$coefficients[2,4], lag_gr_kw$p.value, n16S_gr_kw$p.value, 
                              summary(chabund_gr_lm)$coefficients[2,4], 
                              summary(startabund_gr_lm)$coefficients[2,4], summary(endabund_gr_lm)$coefficients[2,4]), 
                            method = "holm", n=6)

growth_asv_padj
```

Death ASV-level:

```{r}
# generation time
g_de_lm <- lm(log(g) ~ Soil, data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(g_de_lm)) # normal
plot(predict(g_de_lm), resid(g_de_lm)) # homoskedastic
summary(g_de_lm)

# start day
lag_de_lm <- lm(start_day ~ Soil, data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(lag_de_lm)) # non-normal
plot(predict(lag_de_lm), resid(g_de_lm)) # homoskedastic
# Use non-parametric test
lag_de_kw <- kruskal.test(start_day ~ Soil, data=growth_death_asv[growth_death_asv$type=="death",])

# n16S
n16S_de_lm <- lm(n16S ~ Soil, data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(n16S_de_lm)) # non-normal even w/log transform
plot(predict(n16S_de_lm), resid(g_de_lm)) # homoskedastic
# Use non-parametric test
n16S_de_kw <- kruskal.test(n16S ~ Soil, data=growth_death_asv[growth_death_asv$type=="death",])

# Change abund
chabund_de_lm <- lm(log(abs(change_abund_corr)) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(chabund_de_lm)) # normal
plot(predict(chabund_de_lm), resid(chabund_de_lm)) # homoskedastic w/ or w/o log transform
summary(chabund_de_lm)

# Start abund
startabund_de_lm <- lm(log(start_abund_corr) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(startabund_de_lm)) # normal
plot(predict(startabund_de_lm), resid(startabund_de_lm)) # homoskedastic
summary(startabund_de_lm)

# End abund
endabund_de_lm <- lm(log(end_abund_corr) ~ Soil, 
                 data=growth_death_asv[growth_death_asv$type=="death",])
hist(resid(endabund_de_lm)) # normal
plot(predict(endabund_de_lm), resid(endabund_de_lm)) # homoskedastic
summary(endabund_de_lm)

# Multiple test correction
death_asv_padj <- p.adjust(c(summary(g_de_lm)$coefficients[2,4], lag_de_kw$p.value, n16S_de_kw$p.value, 
                              summary(chabund_de_lm)$coefficients[2,4], 
                              summary(startabund_de_lm)$coefficients[2,4], summary(endabund_de_lm)$coefficients[2,4]), 
                            method = "holm", n=6)

death_asv_padj
```

## Metric vs metric

Based largely on results from previously experiments. Want to see if they replicate under different conditions.          

Growth change abund vs death change abund:

```{r}
growth_death_asv %>% 
  select(Soil, ASV, type, change_abund_corr) %>% 
  pivot_wider(names_from=type, values_from=change_abund_corr) %>% 
  filter(!(is.na(death) | is.na(growth))) %>% # only taxa with a growth and death estimate
  ggplot(aes(x=log(growth), y=log(abs(death)))) +
  geom_point() +
  geom_smooth(method="lm", linetype=2, color="black") +
  labs(x="ln deltaNg", y="ln deltaNd") +
  #scale_shape_manual(values=c(1, 19)) +
  facet_wrap(~Soil, scales="free") +
  theme_test() +
  theme()
```

Growth rate vs change abund:

```{r}
growth_death_asv %>%
  select(Soil, ASV, type, g, change_abund_corr) %>% 
  filter(type=="growth") %>% 
  ggplot(aes(x=log(g), y=log(change_abund_corr), color=Soil)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x="ln generation time", y="ln deltaNg") +
  theme_test()
```

Death rate vs change abund:

```{r}
growth_death_asv %>%
  select(Soil, ASV, type, g, change_abund_corr) %>% 
  filter(type=="death") %>% 
  ggplot(aes(x=log(g), y=log(abs(change_abund_corr)), color=Soil)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x="ln halving time", y="ln deltaNd") +
  theme_test()
```

### Statistics

Pearson correlations

DeltaNg vs deltaNd:

```{r}
# Reformat
growth_death_deltaN <- growth_death_asv %>% 
  select(Soil, ASV, type, change_abund_corr) %>% 
  pivot_wider(names_from=type, values_from=change_abund_corr) %>% 
  filter(!(is.na(death) | is.na(growth)))

# Correlate
# Ag
cor.test(log(growth_death_deltaN[growth_death_deltaN$Soil=="Ag",]$growth), log(abs(growth_death_deltaN[growth_death_deltaN$Soil=="Ag",]$death)), method="pearson")
# Meadow
cor.test(log(growth_death_deltaN[growth_death_deltaN$Soil=="Meadow",]$growth), log(abs(growth_death_deltaN[growth_death_deltaN$Soil=="Meadow",]$death)), method="pearson")
```

G vs deltaNg:

```{r}
# Reformat
growth_gdeltaNg <- growth_death_asv %>%
  select(Soil, ASV, type, g, change_abund_corr) %>% 
  filter(type=="growth")

# Correlate
# Ag
cor.test(log(growth_gdeltaNg[growth_gdeltaNg$Soil=="Ag",]$g), log(growth_gdeltaNg[growth_gdeltaNg$Soil=="Ag",]$change_abund_corr), method="pearson")
# Meadow
cor.test(log(growth_gdeltaNg[growth_gdeltaNg$Soil=="Meadow",]$g), log(growth_gdeltaNg[growth_gdeltaNg$Soil=="Meadow",]$change_abund_corr), method="pearson")
```

h vs deltaNd:

```{r}
# Reformat
growth_hdeltaNd <- growth_death_asv %>%
  select(Soil, ASV, type, g, change_abund_corr) %>% 
  filter(type=="death")

# Correlate
# Ag
cor.test(log(growth_hdeltaNd[growth_hdeltaNd$Soil=="Ag",]$g), log(abs(growth_hdeltaNd[growth_hdeltaNd$Soil=="Ag",]$change_abund_corr)), method="pearson")
# Meadow
cor.test(log(growth_hdeltaNd[growth_hdeltaNd$Soil=="Meadow",]$g), log(abs(growth_hdeltaNd[growth_hdeltaNd$Soil=="Meadow",]$change_abund_corr)), method="pearson")
```

## Soil ASV overlap

```{r}
ag_asvs <- growth %>% filter(Soil=="Ag") %>% select(ASV) %>% unique() %>% as.list()
mead_asvs <- growth %>% filter(Soil=="Meadow") %>% select(ASV) %>% unique() %>% as.list()

# Number of ASVs
length(ag_asvs$ASV)
length(mead_asvs$ASV)
length(intersect(ag_asvs$ASV, mead_asvs$ASV))

# Proportions
length(intersect(ag_asvs$ASV, mead_asvs$ASV))/length(ag_asvs$ASV)
length(intersect(ag_asvs$ASV, mead_asvs$ASV))/length(mead_asvs$ASV)
```


# SCRAP

Death:

# Gaussian mixture modeling

```{r}
library(mclust)
set.seed(3)
```

## Growth k

```{r}
# Agricultural

# Visualize
ggplot(data=growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="growth",],
       aes(log(k))) +
  geom_density() +
  theme_test()

# Subset data
k_grag <- growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="growth",]$k

# Fit univariate gaussian mixture models
agk_mclust <- densityMclust(log(k_grag))
summary(agk_mclust)
plot(agk_mclust$BIC)
agk_mclust$parameters
```

```{r}
# Meadow

# Visualize
ggplot(data=growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="growth",],
       aes(log(k))) +
  geom_density() +
  theme_test()

# Subset data
k_grmd <- growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="growth",]$k

# Fit univariate gaussian mixture models
agk_mclust <- densityMclust(log(k_grmd))
summary(agk_mclust)
plot(agk_mclust$BIC)
agk_mclust$parameters
```

## Death k

Fit univariate gaussian mixture models:

# PIN: this seems weird?? look into later
```{r}
# Agricultural

# Visualize
ggplot(data=growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="death",],
       aes(log(k))) +
  geom_density() +
  theme_test()

# Subset data
k_deag <- growth_death_asv[growth_death_asv$Soil=="Ag" & growth_death_asv$type=="death",]$k

# Fit univariate gaussian mixture models
agd_mclust <- densityMclust(log(k_deag))
summary(agd_mclust)
plot(agd_mclust$BIC)
agd_mclust$parameters
```

```{r}
# Meadow

# Visualize
ggplot(data=growth_death_asv[growth_death_asv$Soil=="Meadow" & growth_death_asv$type=="death",],
       aes(log(k))) +
  geom_density() +
  theme_test()

# Subset data
k_demd <- growth_death_asv[growth_death_asv$Soil=="Meadow" & growth_death_asv$type=="death",]$k

# Fit univariate gaussian mixture models
mdd_mclust <- densityMclust(log(abs(k_demd)))
summary(mdd_mclust)
plot(mdd_mclust$BIC)
mdd_mclust$parameters
```

